European or American? A re-examination of the transatlantic influence over the IMF

### 2025-09-24 manuscript/manuscript.qmd
Refine the para above

These findings have several important implications for our understanding of international financial institutions and global economic governance more broadly. Theoretically, these results complicate conventional narratives of uncontested US hegemony in the IMF. While the United States remains the most powerful single actor within the Bretton Woods system, its influence is concentrated in shaping conditionality rather than in determining access to IMF resources. By contrast, European powers collectively wield substantial institutional authority over lending volumes and approvals, reflecting the Fund’s historical leadership structure epitomised by the “gentlemen’s agreement.” Hegemony within the IMF, therefore, is fragmented rather than monolithic, with American primacy tempered by European institutional strength.

>>>>

These findings have several important implications for our understanding of international financial institutions and global economic governance. Theoretically, they complicate conventional narratives of uncontested US hegemony in the IMF. While the United States remains the most powerful single actor within the Bretton Woods system, its influence is concentrated in shaping the *policy content* of lending arrangements—determining what borrowers must do to receive and maintain access to funds—rather than in deciding who receives assistance. By contrast, European powers collectively wield substantial institutional authority over *resource allocation*, influencing which countries obtain loans and in what amounts. This pattern, rooted in the Fund’s historical leadership structure epitomised by the “gentlemen’s agreement,” suggests that hegemony within the IMF is fragmented rather than monolithic, with American primacy tempered by European institutional strength.
